Eco-evolutionary dynamics of finite populations from first principles

### 2023-02-14 ind_files/06_unification_discussion.tex
Cite Rice for discrete time version of stoch price eqn

&- \frac{1}{K}\underbrace{\frac{1}{N_{K}(t)}\left[(\tau_i(\mathbf{x}) - \overline{\tau})p_i + \mu\left\{Q_i(\mathbf{p}) - p_i\left(\sum\limits_{j=1}^{m}Q_j(\mathbf{p})\right)\right\}\right]dt}_{\substack{\text{Directional noise-induced effect: selection-mutation balance}\\\text{for lower turnover rates}}}\\
&+ \frac{1}{\sqrt{K}N_{K}(t)}\underbrace{\left[\left(A^{+}_{i}\right)^{1/2}dB^{(i)}_t - p_i\sum\limits_{j=1}^{m}\left(A^{+}_{j}\right)^{1/2}dB^{(j)}_t\right]}_{\substack{\text{Non-directional noise-induced effect}\\\text{due to stochastic fluctuations}}}

@@

where all covariances are understood in the statistical sense (Note that since $w_i$, $\tau_i$, $\overline{w}$, and $\overline{\tau}$ are stochastic processes depending on $\mathbf{p}$, the terms $\textrm{Cov}(w,f)$ and $\textrm{Cov}(\tau,f)$ are themselves stochastic processes). Taking $K \to \infty$ in equation \eqref{nD_stochastic_Price} recovers the standard Price equation as the infinite population limit (either \eqref{nD_Price_time_dependent} or \eqref{nD_Price} based on whether $f_i$ varies with time). We saw in chapter \ref{chap_infD_processes} using field equations that very similar relations hold for quantitative traits, and indeed, measure-theoretic tools have recently been used to rigorously show that an infinite-dimensional version of \eqref{nD_stochastic_Price} holds for one-dimensional quantitative traits when $f$ is the trait value and $b(x|\psi) \pm d(x|\psi)$ are Gaussian \citep{week_white_2021}. Equations \eqref{nD_stochastic_RM} and \eqref{nD_stochastic_Price} are thus the `fundamental' equations for the evolution of finite populations, with the replicator-mutator and Price equations as their respective infinite population limits.\\

@@

\mathbb{E}\left[\frac{d\overline{\tau}}{dt}\right] = \underbrace{ \vphantom{ \mathbb{E}\left[\frac{\textrm{Var}(\tau)}{KN_K(t)}\right] } \mathbb{E}\left[\textrm{Var}(b^{\textrm{(ind)}}) - \textrm{Var}(d^{\textrm{(ind)}}\right]}_{\substack{\text{Classical selection} \\ \text{On life-history}}} \ - \ \underbrace{ \mathbb{E}\left[\frac{\textrm{Var}(\tau)}{KN_K(t)}\right]}_{\substack{\text{Noise-induced selection} \\ \text{On life-history}}} \ + \ \underbrace{ \vphantom{ \mathbb{E}\left[\frac{\textrm{Var}(\tau)}{KN_K(t)}\right] } \sum\limits_{i=1}^{m}\mathbb{E}\left[p_i\frac{\partial \tau_i}{\partial t}\right]}_{\substack{\text{Short-term (ecological)} \\ \text{feedbacks to $\tau_i$}}}

>>>>

&- \frac{1}{K}\underbrace{\frac{1}{N_{K}(t)}\left[(\tau_i(\mathbf{x}) - \overline{\tau})p_i + \mu\left\{Q_i(\mathbf{p}) - p_i\left(\sum\limits_{j=1}^{m}Q_j(\mathbf{p})\right)\right\}\right]dt}_{\substack{\text{Directional noise-induced effects: selection-mutation balance}\\\text{for lower turnover rates}}}\\
&+ \frac{1}{\sqrt{K}N_{K}(t)}\underbrace{\left[\left(A^{+}_{i}\right)^{1/2}dB^{(i)}_t - p_i\sum\limits_{j=1}^{m}\left(A^{+}_{j}\right)^{1/2}dB^{(j)}_t\right]}_{\substack{\text{Non-directional noise-induced effects}\\\text{due to stochastic fluctuations}}}

@@

where all covariances are understood in the statistical sense (Note that since $w_i$, $\tau_i$, $\overline{w}$, and $\overline{\tau}$ are stochastic processes depending on $\mathbf{p}$, the terms $\textrm{Cov}(w,f)$ and $\textrm{Cov}(\tau,f)$ are themselves stochastic processes). Taking $K \to \infty$ in equation \eqref{nD_stochastic_Price} recovers the standard Price equation as the infinite population limit (either \eqref{nD_Price_time_dependent} or \eqref{nD_Price} based on whether $f_i$ varies with time). We saw in chapter \ref{chap_infD_processes} using field equations that very similar relations hold for quantitative traits, and indeed, measure-theoretic tools have recently been used to rigorously show that an infinite-dimensional version of \eqref{nD_stochastic_Price} holds for one-dimensional quantitative traits when $f$ is the trait value and $b(x|\psi) \pm d(x|\psi)$ are Gaussian \citep{week_white_2021}. Equations \eqref{nD_stochastic_RM} and \eqref{nD_stochastic_Price} are thus the `fundamental' equations for the evolution of finite populations, with the replicator-mutator and Price equations as their respective infinite population limits (also see \citep{rice_universal_2020} for a stochastic Price equation in a discrete-time setting).\\

@@

\mathbb{E}\left[\frac{d\overline{\tau}}{dt}\right] = \underbrace{ \vphantom{ \mathbb{E}\left[\frac{\textrm{Var}(\tau)}{KN_K(t)}\right] } \mathbb{E}\left[\textrm{Var}(b^{\textrm{(ind)}}) - \textrm{Var}(d^{\textrm{(ind)}}\right]}_{\substack{\text{Classical selection} \\ \text{effects}}} \ - \ \underbrace{ \mathbb{E}\left[\frac{\textrm{Var}(\tau)}{KN_K(t)}\right]}_{\substack{\text{Noise-induced selection} \\ \text{effects}}} \ + \ \underbrace{ \vphantom{ \mathbb{E}\left[\frac{\textrm{Var}(\tau)}{KN_K(t)}\right] } \sum\limits_{i=1}^{m}\mathbb{E}\left[p_i\frac{\partial \tau_i}{\partial t}\right]}_{\substack{\text{Short-term (ecological)} \\ \text{feedbacks to $\tau_i$}}}
